---
title: "Communicable disease snapshot"
subtitle: "`r params$county` County"
date: today
format:
  html:
    embed-resources: true
    theme: cosmo
    mainfont: "Open Sans"
execute:
  echo: false
  warning: false
params:
  county: "Kern"
  year: 2023
---

```{r}
#| label: setup
# On Posit Connect, scheduled renders have no viewer, so this report uses a
# *service account* Databricks integration (Access > Integrations). The
# code is identical to the interactive scripts: odbc::databricks() (odbc
# 1.7.0 or later) picks up the service account token from connectcreds.
library(DBI)
library(dplyr, warn.conflicts = FALSE)
library(dbplyr, warn.conflicts = FALSE)
library(ggplot2)
library(gt)
library(connectcreds)

con <- dbConnect(
  odbc::databricks(),
  httpPath = Sys.getenv("DATABRICKS_HTTP_PATH"),
  bigint = "numeric"
)

cases <- tbl(con, I("public_health_demo.surveillance.infectious_disease_cases")) |>
  filter(sex == "Total")

county_year <- cases |>
  filter(county == !!params$county, year == !!params$year) |>
  select(disease, cases, rate_per_100k, rate_unstable)

state_year <- cases |>
  filter(is_statewide, year == !!params$year) |>
  select(disease, state_rate = rate_per_100k)

snapshot <- county_year |>
  left_join(state_year, by = "disease") |>
  filter(is.na(cases) | cases > 0) |>
  arrange(desc(rate_per_100k)) |>
  collect()

top <- slice_head(snapshot, n = 1)
```

In `r params$year`, `r nrow(snapshot)` reportable diseases had at least one
case among `r params$county` County residents. The highest incidence was
**`r top$disease`** at **`r round(top$rate_per_100k, 1)` per 100,000**,
compared with `r round(top$state_rate, 1)` statewide.

## Diseases reported in `r params$year`

```{r}
#| label: table
snapshot |>
  mutate(ratio = rate_per_100k / state_rate) |>
  relocate(rate_unstable, .after = ratio) |>
  gt() |>
  cols_label(
    disease = "Disease", cases = "Cases", rate_per_100k = "County rate",
    state_rate = "State rate", ratio = "County / state", rate_unstable = "Unstable"
  ) |>
  fmt_number(c(rate_per_100k, state_rate), decimals = 1) |>
  fmt_number(ratio, decimals = 1, pattern = "{x}x") |>
  fmt_number(cases, decimals = 0) |>
  sub_missing(c(cases, rate_per_100k, ratio), missing_text = "Suppressed") |>
  text_transform(
    locations = cells_body(rate_unstable),
    fn = function(x) ifelse(x == "TRUE", "Yes", "")
  ) |>
  tab_source_note("Counts of 1-10 and their rates are suppressed. Rates per 100,000. Unstable rates have a relative standard error of 23% or more.")
```

## Trend for `r top$disease`

```{r}
#| label: trend
#| fig-height: 4
trend <- cases |>
  filter(disease == !!top$disease, county %in% !!c(params$county, "California")) |>
  select(county, year, rate_per_100k) |>
  collect() |>
  mutate(county = if_else(county == "California", "California (statewide)", county))

ggplot(trend, aes(year, rate_per_100k, color = county)) +
  geom_line(linewidth = 1) +
  geom_point() +
  scale_color_manual(values = c("#EE6331", "#447099")) +
  labs(x = NULL, y = "Rate per 100,000", color = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")
```

```{r}
#| label: cleanup
dbDisconnect(con)
```

*Source: CDPH Infectious Diseases Branch, CHHS Open Data Portal. Generated from Databricks with Posit.*
